---
title: "Load and Reuse a BERTopic Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Load and Reuse a BERTopic Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

This vignette shows how to load a previously saved BERTopic model in a new session and reuse the
extras stored alongside it. Set `eval = TRUE` for the chunks you want to run.

## Environment setup

```{r}
library(reticulate)
library(bertopicr)
library(readr)
library(dplyr)

# macOS: if reticulate fails to load Python libraries, run once per session.
if (identical(Sys.info()[["sysname"]], "Darwin")) {
  bertopicr::configure_macos_homebrew_zlib()
}

# Point reticulate at the same Python environment used for training.
reticulate::use_virtualenv("c:/path/to/your/venv", required = TRUE)

reticulate::py_config()
reticulate::py_available()
reticulate::py_run_string(code = "import torch
print(torch.cuda.is_available())") # if GPU is available then TRUE else FALSE)

```

## Load the model bundle

```{r}
loaded <- load_bertopic_model("topic_model") # set the location of the model!
model <- loaded$model
extras <- loaded$extras
```

## Load data for inspection

```{r}
sample_path <- system.file("extdata", "spiegel_sample.rds", package = "bertopicr")
df <- read_rds(sample_path)
docs <- df |> pull(text_clean)
```

## Create tables from the loaded model

```{r}
doc_info <- get_document_info_df(model = model, texts = docs)
topic_info <- get_topic_info_df(model = model)
topics_df <- get_topics_df(model = model)
```

## Use extras and visualizations

```{r}
visualize_barchart(model = model, filename = "barchart_demo")
visualize_distribution(
  model = model,
  text_id = 1,
  probabilities = extras$probabilities,
  filename = "vis_topic_dist_demo"
)
visualize_heatmap(model = model, filename = "vis_heat_demo")
visualize_topics(model = model, filename = "dist_map_demo")

```


```{r}
visualize_documents(model = model, docs, reduced_embeddings = extras$reduced_embeddings_2d)
visualize_documents_2d(model = model, docs, reduced_embeddings = extras$reduced_embeddings_2d)
visualize_documents_3d(model = model, docs, reduced_embeddings = extras$reduced_embeddings_3d)

```

The following visualizations work only if *topics_over_time* and *topics_per_class* were defined after model training or within the `train_bertopic_model()` function.

```{r, eval=FALSE}
visualize_topics_over_time(model = model, topics_over_time_model = extras$topics_over_time)
visualize_topics_per_class(model, extras$topics_per_class, auto_open = FALSE)

```
